---
title: "T1D Lipidome"
author: Tommi Suvitaival, tommi.raimo.leo.suvitaival@regionh.dk, Steno Diabetes Center
  Copenhagen
date: "`r Sys.Date()`"
output: github_document
---

```{r setup, include=FALSE}

knitr::opts_chunk$set( echo = TRUE )

knitr::opts_knit$set( root.dir = normalizePath( '../' ) )
              
```

# Preparations

## Load Data

```{r Load}

data.loaded <-
  read.table(
    file = "data-updated/T1D-lipidome-database-updated.tsv",
    check.names = FALSE,
    header = TRUE,
    na = "",
    sep = "\t",
    stringsAsFactors = FALSE
  )

```

## Create Additional Columns

```{r Format, message = FALSE, warning = FALSE}

data <- tibble::as_tibble( x = data.loaded  )

tmp <-
  apply(
    X = is.na( data ),
    MAR = 1,
    FUN = all
  )

data <- data[ !tmp, ]

data$"Authors.list" <-
  stringr::str_split(
    string = data$"Authors",
    pattern = ", "
  )

data$"Author.Year.Journal" <-
  paste0(
    data$"Author",
    " (",
    data$"Year",
    ") ",
    data$"Journal"
  )

data$"N.Limiting" <-
  apply(
    X = as.matrix( data[ , c( "N Case", "N Control" ) ] ),
    MAR = 1,
    FUN = 
      function( x ) {
        
        tmp <- min( na.omit( x ) )
        
        if ( is.infinite( tmp ) ) {
          
          tmp <- NA
          
        }
        
        return( tmp )
        
      }
  )

data$"Name" <-
  stringr::str_replace(
    string = data$"Name",
    pattern = "Lyso",
    replacement = "L"
  )

```

## Number of Entries

```{r}

dim( data[ !is.na( data$"N.Limiting" ), ] )

dim( data[ !is.na( data$"N.Limiting" ) & is.na( data$"Time from diagnosis" ), ] )

dim( data[ !is.na( data$"N.Limiting" ) & !is.na( data$"Time from diagnosis" ), ] )

```

\newpage

# Progressors to T1D v. Control

```{r Papers-Cor}

names.compounds <- sort( unique( data$"Name" ) )

data.plot <- data[ data$"Contrast" == "PT1D/CTR", ]

data.plot <- data.plot[ !is.na( data.plot$"Significance" ), ]

age <- c( "Birth", "1 y" )

data.plot <- data.plot[ data.plot$Age %in% age, ]

data.plot.wide <-
  tidyr::pivot_wider(
    data = data.plot,
    names_from = "Name",
    values_from = "Sign"
  )

data.plot.wide <-
  data.plot.wide[ is.na( data.plot.wide$"Comment" ), ]

rownames( data.plot.wide ) <-
  data.plot.wide$"Author.Year.Journal"

cor.papers <-
  cor(
    x = t( data.plot.wide[ , colnames( data.plot.wide ) %in% names.compounds ] ),
    use = "pairwise.complete.obs"
  )

table.paper.by.name <- table( data$"Author.Year.Journal", data$"Name" )

cor.papers <- cor( t( table.paper.by.name ) )

cov.papers <- cov( t( table.paper.by.name ) )

```

\newpage

# Consensus of Aberrations in the Lipidome Before the Onset of T1D

## Prepare Data

### Map Lipid Names

```{r Map-Names}

data.plot <- data[ data$"Contrast" == "PT1D/CTR", ]

data.plot <- data.plot[ !is.na( data.plot$"Significance" ), ]

names.lipids.unique <- unique( data.plot$"Name" )

names.mapping <- lipidomeR::map_lipid_names( x = names.lipids.unique )

names.mapping[ which( names.mapping$"Class" == "PC-O/" ), "Class" ] <-
  "PC-O/P"

```

### Compute Consensus Aberrations

```{r Compute Consensus}

data.plot <-
  merge(
    x = data.plot,
    y = names.mapping,
    by = "Name",
    all = TRUE
  )

data.grouped <-
  dplyr::group_by(
    .data = data.plot,
    Name.simple,
    Author.Year.Journal
  )

data.grouped <-
  dplyr::summarise(
    .data = data.grouped,
    Sign.median = median( Sign )
  )

data.summed <-
  dplyr::group_by(
    .data = data.grouped,
    Name.simple
  )

data.summed <-
  dplyr::summarise(
    .data = data.summed,
    Consensus_Effect = sum( Sign.median )
  )

data.summed <-
  merge(
    x = data.summed,
    y = names.mapping,
    all.x = TRUE,
    all.y = FALSE
  )

```

\newpage

## Heatmap of Consensus Aberration in Progression to T1D

```{r Consensus-Sign-Heatmap, fig.width = 6, fig.height = 6, dpi = 600, dev = c( "svg" )}

tmp <- c( "Name", "Consensus_Effect" )

tmp2 <- colnames( data.summed )[ !( colnames( data.summed ) %in% tmp ) ]
tmp2 <- c( "Name", tmp2 )

figure <-
  lipidomeR::heatmap_lipidome(
    x = data.summed[ , tmp ],
    names.mapping = data.summed[ , tmp2 ],
    class.facet = "wrap",
    x.names = "Name",
    melt.value.name = "Consensus_Effect",
    fill.limits = 
      c( -1, 1 ) * 
      max( 
        x = abs( data.summed$"Consensus_Effect" ),
        na.rm = TRUE
      ),
    fill.midpoint = 0,
    scales = "free",
    space = "free",
    axis.x.carbons = FALSE
  )

# print( figure )

figure.B <-
  figure +
  ggplot2::xlab( 
    label = "Lipid degree of unsaturation\n(number of fatty-acid double bonds)"
  ) +
  ggplot2::ylab(
    label = "Lipid size\n(number of fatty-acid carbon atoms)"
  )

print( figure.B )

```

## Counts of Consensus-Aberrated Lipids

```{r}

tmp <- unique( data.plot$"Name.simple" )

length( tmp ) 

tmp <-
  stringr::str_split_fixed(
    string = tmp,
    pattern = "\\(",
    n = 2
  )

tmp <-
  stringr::str_replace(
    string = tmp[ , 1 ],
    pattern = "O\\/$",
    replacement = "O/P"
  )

table( tmp )

sort( table( tmp ) )

table( data.plot$"Name.simple" )

table( table( data.plot$"Name.simple" ) > 1 )

table( table( data.grouped$"Name.simple" ) )

tmp <- table( data.grouped$"Name.simple" )

tmp <- tmp[ tmp > 1 ]

length( tmp )

```

\newpage

# Similarity Network of Publications Based on the Lipidome

## Compute Similarity

```{r Data-Network}

data.plot <-
  data[
    !is.na( data$"Significance" ) &
      ( !is.na( data$"N Case" ) | !is.na( data$"N Control" )
      ),
    ]

data.plot <- data.plot[ !grepl( x = data.plot$"Name", pattern = "Oxidized" ), ]

names.lipids.unique <- unique( data.plot$"Name" )

names.mapping <- lipidomeR::map_lipid_names( x = names.lipids.unique )

names.mapping[ which( names.mapping$"Class" == "PC-O/" ), "Class" ] <-
  "PC-O/P"

data.plot <-
  merge(
    x = data.plot,
    y = names.mapping,
    by = "Name",
    all = TRUE
  )

data.grouped <-
  dplyr::group_by(
    .data = data.plot,
    Name.simple,
    Author.Year.Journal
  )

data.grouped <-
  dplyr::summarise(
    .data = data.grouped,
    Sign.median = median( Sign )
  )

data.wide <-
  tidyr::pivot_wider(
    data = data.grouped,
    names_from = "Name.simple",
    values_from = "Sign.median"
  )

data.wide[ is.na( data.wide ) ] <- 0

cor.data <-
  cor(
    x = t( data.wide[ , -1 ] ),
    use = "pairwise.complete.obs",
    method = "spearman"
  )

colnames( cor.data ) <- 
  rownames( cor.data ) <- 
  data.wide$"Author.Year.Journal"

agreement <-
  array( dim = nrow( data.wide ) * c( 1, 1 ) )

rownames( agreement ) <-
  colnames( agreement ) <-
  data.wide$"Author.Year.Journal"

disagreement <-
  jaccard <-
  jaccard.identity <-
  jaccard.disagreement <-
  agreement

jaccard.others <- rep( x = NA, times = nrow( data.wide ) )
names( jaccard.others ) <- rownames( agreement )

for ( i in 1:( nrow( data.wide ) - 1 ) ) {

  tmp.i <- unlist( data.wide[ i, -1 ] )
      
  for ( j in ( i + 1 ):nrow( data.wide ) ) {
    
    tmp.j <- unlist( data.wide[ j, -1 ] )
    
    agreement[ i, j ] <-
      agreement[ j, i ] <-
      sum( 
        tmp.i != 0 &
          sign( tmp.i ) == sign( tmp.j )
      )
    
    disagreement[ i, j ] <-
      disagreement[ j, i ] <-
      sum( 
        tmp.i != 0 &
          tmp.j != 0 &
          sign( tmp.i ) != sign( tmp.j )
      )
    
    jaccard[ i, j ] <-
      jaccard[ j, i ] <-
      sum(
        tmp.i != 0 &
          tmp.j != 0
      ) /
      sum(
        tmp.i != 0 |
          tmp.j != 0
      )
    
    jaccard.identity[ i, j ] <-
      jaccard.identity[ j, i ] <-
      sum(
        tmp.i != 0 &
          sign( tmp.i ) == sign( tmp.j )
      ) /
      sum(
        tmp.i != 0 |
          tmp.j != 0
      )
    
    jaccard.disagreement[ i, j ] <-
      jaccard.disagreement[ j, i ] <-
      sum(
        tmp.i != 0 &
          tmp.j != 0 &
          sign( tmp.i ) != sign( tmp.j )
      ) /
      sum(
        tmp.i != 0 |
          tmp.j != 0
      )
    
  }
    
}

for ( i in 1:nrow( data.wide ) ) {
  
  tmp.i <- unlist( data.wide[ i, -1 ] )
      
  jaccard.others[ i ] <-
    sum(
      tmp.i !=0 &
        apply(
          X = data.wide[ -i, -1 ] != 0,
          MAR = 2,
          FUN = any
        )
    ) /
    sum(
      apply(
        X = data.wide[ , -1 ] != 0,
        MAR = 2,
        FUN = any
      )
    )
  
}

```

## Network Figure

```{r Papers-Network-Mapped, fig.width = 9, fig.height = 9, dpi = 600, dev = c( "svg" )}

is.post.diagnosis <- 
  !is.na(
    unlist( data[ 
      match( 
        x = rownames( cor.data ),
        table = data$"Author.Year.Journal"
      ),
      "Time from diagnosis"
      ]
    )
  )

names( is.post.diagnosis ) <- rownames( cor.data )

is.multiple.testing.corrected <-
    grepl(
      x = 
        unlist( data[ 
          match( 
            x = rownames( cor.data ),
            table = data$"Author.Year.Journal"
          ),
          "Significance"
          ]
        ),
      pattern = "adj"
  )

names( is.multiple.testing.corrected ) <- rownames( cor.data )

edge.labels <- as.character( round( x = jaccard.identity * 100 ) )

tmp <- as.character( round( x = jaccard.disagreement* 100 ) )

edge.labels <-
  paste(
    edge.labels,
    tmp,
    sep =","
  )

edge.labels[ edge.labels == "NA; NA" ] <- NA

edge.labels <- array( data = edge.labels, dim = dim( jaccard.identity ) )

jaccard.with.sign <- jaccard

tmp <- jaccard.disagreement > jaccard.identity
jaccard.with.sign[ which( tmp ) ] <- jaccard.with.sign[ which( tmp ) ] * -1

labels <-
  stringr::str_wrap(
    string = colnames( jaccard.with.sign ),
    width = 10
  )

```

* Nodes: Studies
* Edge width: Similarity in reported lipids between two studies
* Edge label: Agreement and disagreement in the sign of the aberration of the lipids (first value: agreement; second value: disagreement)
* Edge color: Majority of agreement (red) or disagreement (blue) in the sign of aberration
* Node size: Average similarity to the other studies
* Node shape: Pre-onset or post-onset study (circle: pre-onset; square: post-onset)
* Node color: Reported significances corrected for multiple testing (gray) or not corrected (white)

```{r Papers-Network-Publication, fig.width = 9, fig.height = 9, dpi = 600, dev = c( "svg" )}

qgraph::qgraph(
  input = jaccard.with.sign,
  diag = FALSE,
  layout = "spring",
  posCol = "red",
  negCol = "blue",
  edge.labels = edge.labels,
  edge.label.margin = 0.02,
  color = 
    ifelse(
      test = is.multiple.testing.corrected,
      yes = "gray",
      no = "white"
    ),
  shape =
    ifelse(
      test = is.post.diagnosis,
      yes = "rectangle",
      no = "circle"
    ),
  labels = labels,
  vsize = 10 * sqrt( jaccard.others ) + 10,
  mar = c( 1, 1, 1, 1 ) * 2
)

```

```{r Agreement-Statistics}

summary( abs( c( jaccard.with.sign ) ) )

rowMeans(
  x = abs( jaccard.with.sign ),
  na.rm = TRUE
)

```

\newpage

# Summary Table of Lipid Classes

```{r Table-Lipid-Classes-Prepare}

data.table <- 
  data.plot[ 
    !is.na( data.plot$"N Case" ) |
      !is.na( data.plot$"N Control" ),
    ]

res <-
  by(
    data = data.table,
    INDICES = data.table$"Class",
    FUN =
      function( x ) {
        tmp <- table( x$"Name.simple", x$"Author.Year.Journal" )
        tmp2 <- apply( X = tmp > 0, MAR = 1, FUN = sum )
        
        tmp3 <- which( tmp2 == max( x = tmp2, na.rm = TRUE ) )
        
        if ( length( tmp3 ) > 1 ) {
          tmp4 <- median( x = x[ tmp3, "N.double.bonds" ], na.rm = TRUE )
          tmp5 <- median( x = x[ tmp3, "N.carbons" ], na.rm = TRUE )

          tmp3 <-
            tmp3[
              which.min(
                ( x[ tmp3, "N.double.bonds" ] - tmp4 )^2 +
                  ( x[ tmp3, "N.carbons" ] - tmp5 )^2
              )
              ]

        }
          
        tmp4 <- which( x$"Name.simple" %in% names( tmp2 )[ tmp3 ] )[ 1 ]
        
        return(
          data.frame(
            x[ tmp4, ]
          )
        )
          
      }
  )

res <- dplyr::bind_rows(
  res,
  .id = "Lipid_Class"
)

tmp <- 
  tapply(
    X = data.table$"Name.simple",
    INDEX = data.table$"Class",
    FUN =
      function( x ) {
        length( unique( x ) )
      }
  )

res$"N.unique" <- tmp[ match( x = names( tmp ), table = res$"Class" ) ]

# View( data.table[ data.table$"Name.simple" %in% res$"Name.simple", ] )

```

```{r Table-Lipid-Classes-Print}

res$"Lipid.Class.Name" <-
  c(
    "Cholesterol-ester",
    "Ceramide",
    "Diacylglycerol",
    "Lyso-phosphatidylcholine",
    "Lyso-phosphatidylethanolamine",
    "Phosphatidylcholine",
    "Alkyl-acyl PC",
    "Phosphatidylethanolamine",
    "Alkyl-acyl PE",
    "Phosphatidylinositol",
    "Plasmenyl-PC",
    "Plasmenyl-PE",
    "Phosphatidylserine",
    "Sphingomyelin",
    "Triacylglycerol"
  )

# res$"Lipid.Class.Name" <-
#   c(
#     "Cholesterol-ester",
#     "Ceramide",
#     "Diacylglycerol",
#     "Lyso-phosphatidylcholine",
#     "Lyso-phosphatidylethanolamine",
#     "Phosphatidylcholine",
#     "Alkyl-acyl PC",
#     "Phosphatidylethanolamine",
#     "Alkyl-acyl PE",
#     "Phosphatidylinositol",
#     "Sphingomyelin",
#     "Triacylglycerol"
#   )

names.columns <- 
  c(
    "Class",
    "Lipid.Class.Name",
    "N.unique",
    "Name.simple"
  )

lipid.classes.table <- res[ , names.columns ]

colnames( lipid.classes.table ) <-
  c(
    "Class abbreviation",
    "Class name",
    "Number of unique species in review",
    "Example species"
  )

# knitr::kable( x = lipid.classes.table )
# knitr::kable( x = tibble::as_tibble( lipid.classes.table ) )

write.table(
  x = lipid.classes.table,
  file = "results-updated/Lipid_Classes.tsv",
  row.names = FALSE,
  sep = "\t"
)

```

\newpage

# Heatmap of Lipids by Publication

```{r Heatmap-Lipids-by-Publications, fig.width = 9, fig.height = 9, dpi = 600, dev = c( "svg" )}

data.heatmap <- as.matrix( data.wide[ , -1 ] )

rownames( data.heatmap ) <- data.wide$"Author.Year.Journal"

rownames( data.heatmap ) <-
  stringr::str_replace(
    string = rownames( data.heatmap ),
    pattern = " ",
    replacement = "\n"
  )

rownames( data.heatmap ) <-
  stringr::str_replace(
    string = rownames( data.heatmap ),
    pattern = " ",
    replacement = "    \n"
  )

rownames( data.heatmap ) <-
  stringr::str_replace(
    string = rownames( data.heatmap ),
    pattern = "\\\n",
    replacement = "        \n"
  )

rownames( data.heatmap ) <-
  stringr::str_replace(
    string = rownames( data.heatmap ),
    pattern = "La        \\\nTorre    \n(2013)",
    replacement = "La Torre        \n(2013)    \n"
  )

rownames( data.heatmap ) <-
  stringr::str_replace(
    string = rownames( data.heatmap ),
    pattern = "Oresic        \\\n\\(2008\\)    \\nJournal of ",
    replacement = "Oresic (2008)        \nJournal of    \n"
  )

colnames( data.heatmap ) <-
  stringr::str_replace(
    string = colnames( data.heatmap ),
    pattern = "O\\/\\(",
    replacement = "O/P("
    
  )

result.heatmap <-
  gplots::heatmap.2(
    x = t( data.heatmap ),
    col = gplots::bluered( n = 3 ),
    trace = "none",
    tracecol = "black",
    srtCol = 45,
    cexCol = 0.2 + 1/log10( nrow( data.heatmap ) ) - 0.6,
    margins = c( 1, 1 ) * 6
  )

```

## Heatmap of Lipids Reported in More than One Study

```{r Heatmap-Lipids-by-Publications-Min-2, fig.width = 9, fig.height = 9, dpi = 600, dev = c( "svg" )}

data.heatmap.min2 <- data.heatmap[ , colSums( data.heatmap != 0 ) > 1 ]

result.heatmap <-
  gplots::heatmap.2(
    x = t( data.heatmap.min2 ),
    col = gplots::bluered( n = 3 ),
    trace = "none",
    tracecol = "black",
    srtCol = 45,
    cexCol = 0.2 + 1/log10( nrow( data.heatmap.min2 ) ) - 0.6,
    margins = c( 1, 1 ) * 6
  )

```

## Counts by Lipid Class

```{r}

dim( data.heatmap.min2 )

tmp <-
  stringr::str_split_fixed(
    string = colnames( data.heatmap.min2 ),
    pattern = "\\(",
    n = 2
  )

sort( table( tmp[ , 1 ] ) )

```

```{r}

# View( t( data.heatmap.min2[ , order( colSums( data.heatmap.min2 != 0 ), decreasing = TRUE ) ] ) )

data.clustered <-
  t( data.heatmap.min2[ result.heatmap$"colInd", result.heatmap$"rowInd" ] )

data.clustered <- data.clustered[ nrow( data.clustered ):1, ]

data.clustered <- data.clustered[ rowSums( data.clustered != 0 ) > 1, ]

names.clusters <- 
  publications.clusters <-
  signs.clusters <-
  list()

for ( i in 1:nrow( data.clustered ) ) {
  
  tmp.i <- names( which( data.clustered[ i, ] != 0 ) )
  
  if ( i == 1 ) {

    names.clusters[[ 1 ]] <- rownames( data.clustered )[ 1 ]
    
    publications.clusters[[ 1 ]] <- tmp.i
    
    signs.clusters[[ 1 ]] <- data.clustered[ 1, ]
        
  } else {
    
    k <- length( names.clusters )
      
    if (
      all(
        data.clustered[ i, ] ==
        signs.clusters[[ k ]] )
    ) {
    
      names.clusters[[ k ]] <- 
        c(
          names.clusters[[ k ]],
          rownames( data.clustered )[ i ]
        )
      
    } else {
      
      names.clusters[[ k+1 ]] <- rownames( data.clustered )[ i ]
      publications.clusters[[ k+1 ]] <- tmp.i
      signs.clusters[[ k+1 ]] <- data.clustered[ i, ]
      
    }
      
  }
  
}

names.clusters <-
  lapply(
    X = names.clusters,
    FUN = sort,
    decreasing = FALSE
  )

signs.clusters.mat <- simplify2array( x = signs.clusters )

signs.clusters.mat <- 
  signs.clusters.mat[ rowSums( signs.clusters.mat !=0 ) > 0, ] 

order.clusters <-
  order( colSums( signs.clusters.mat != 0 ), decreasing = TRUE )

order.clusters <- 1:ncol( signs.clusters.mat )

mode( signs.clusters.mat ) <- "character"

signs.clusters.mat[ signs.clusters.mat == "-1" ] <- "-"
signs.clusters.mat[ signs.clusters.mat == "1" ] <- "+"
signs.clusters.mat[ signs.clusters.mat == "0" ] <- ""

clusters <-
  data.frame(
    t( signs.clusters.mat ),
    check.names = FALSE
  )

clusters.printed <- 
  data.frame(
    Lipids = 
      sapply(
        X = names.clusters,
        FUN = paste,
        collapse = ", "
      ),
    clusters,
    check.names = FALSE
)

clusters.printed <- clusters.printed[ order.clusters, ]
    
colnames( clusters.printed )[ -1 ]

colnames( clusters.printed ) <-
  stringr::str_replace_all(
    string = colnames( clusters.printed ),
    pattern = "\\s+\\n",
    replacement = " "
  )

colnames( clusters.printed )[ -1 ] <-
  paste(
    LETTERS[ 1:( ncol( clusters.printed ) - 1 ) ],
    colnames( clusters.printed[ -1 ] )
  )

clusters.printed$"Total" <- rowSums( clusters.printed[ , -1 ] != "" )

# View( x = clusters.printed )
# knitr::kable( x = clusters.printed )
knitr::kable( x = tibble::as_tibble( clusters.printed ) )

write.table(
  x = clusters.printed,
  file = "results-updated/Lipids_by_Publication.tsv",
  row.names = FALSE,
  sep = "\t"
)

```

\newpage

# Similarity Network of Publications Based on the Author List

* Node size: Limiting sample size in the study
* Node color: Country of the cohort
* Edge width: Similarity in the author list (%)

```{r Author-Network, fig.width = 9, fig.height = 9, dpi = 600, dev = c( "svg" ), message = FALSE, warning = FALSE}

data.authors <-
  data[
    match(
      x = rownames( jaccard ),
      table = data$"Author.Year.Journal"
    ), ]

color <-
  factor(
    x = data.authors$"Population",
    levels = sort( unique( data.authors$"Population" ) ),
    labels = c( "red", "white", "black", "yellow", "blue" )
  )

color.labels <-
  factor(
    x = color,
    labels = c( "white", "blue", "yellow", "blue", "white" )
  )

size.nodes <-
  tapply(
    X = 
      apply(
        X = as.matrix( data[ , c( "N Case", "N Control" ) ] ),
        MAR = 1,
        FUN = 
          function( x ) {
            
            tmp <- min( na.omit( x ) )
            
            if ( is.infinite( tmp ) ) {
              
              tmp <- NA
              
            }
            
            return( tmp )
            
          }
      ),
    INDEX = data$"Author.Year.Journal",
    FUN = median,
    na.rm = TRUE
  )

size.nodes <- size.nodes[ data.authors$"Author.Year.Journal" ]

jaccard.authors <- array( dim = nrow( data.authors ) * c( 1, 1 ) )

for ( i in 1:( nrow( data.authors ) - 1 ) ) {
  
  for ( j in ( i + 1 ):nrow( data.authors ) ) {
  
    jaccard.authors[ i, j ] <-
      jaccard.authors[ j, i ] <-
      sum(
        data.authors$"Authors.list"[[ i ]] %in% 
          data.authors$"Authors.list"[[ j ]]
      ) / 
      length( 
        unique(
          c(
            data.authors$"Authors.list"[[ i ]],
            data.authors$"Authors.list"[[ j ]]
          )
        )
      )

  }
      
}

rownames( jaccard.authors ) <-
  colnames( jaccard.authors ) <-
  data.authors$"Author.Year.Journal"

independence.authors <- NULL

for ( i in 1:nrow( data.authors ) ) {
  
  independence.authors[ i ] <- 
    1 -
    sum(
      data.authors$"Authors.list"[[ i ]] %in% 
        unique( unlist( data.authors[ -i, "Authors.list" ] ) )
    ) / 
    length( data.authors$"Authors.list"[[ i ]] )

}

names( independence.authors ) <- data.authors$"Author.Year.Journal"

edge.labels <- as.character( round( x = jaccard.authors * 100 ) )

edge.labels <-
  paste(
    edge.labels,
    "%"
  )

edge.labels[ edge.labels == "NA %" ] <- NA

edge.labels <- array( data = edge.labels, dim = dim( jaccard.authors ) )

qgraph::qgraph(
  input = jaccard.authors,
  diag = FALSE,
  layout = "spring",
  posCol = "gray",
  edge.labels = edge.labels,
  edge.label.margin = 0.02,
  edge.label.color = "black",
  fade = FALSE,
  shape = 
    ifelse(
      test = is.post.diagnosis,
      yes = "rectangle",
      no = "circle"
    ),
  label.color = as.character( color.labels ),
  color = as.character( color ),
  labels = labels,
  vsize = sqrt( size.nodes ) + 5
)

```

# Summary of Studies

```{r Table1}

data.table1 <- data[ !is.na( data$"N.Limiting" ), ]

data.table1$"N.Limiting.Median" <-
  apply(
    X = data.table1,
    MAR = 1,
    FUN =
      function( x ) {
        tmp <-
          which( data.table1$"Author.Year.Journal" == x$"Author.Year.Journal" )
        median(
          x = unlist( data.table1[ tmp, "N.Limiting" ] ),
          na.rm = TRUE
        )
      }
  )

data.table1$"N.Lipids.Reported" <-
  apply(
    X = data.table1,
    MAR = 1,
    FUN =
      function( x ) {
        tmp <-
          which( data.table1$"Author.Year.Journal" == x$"Author.Year.Journal" )
        length(
          unique(
            unlist( data.table1[ tmp, "Name" ] )
          )
        )
      }
  )

data.table1 <-
  by(
    data = data.table1,
    INDICES = data.table1$"Author.Year.Journal",
    FUN = function( x ) {
      apply(
        X = x,
        MAR = 2,
        FUN =
          function( x ) {
            
            y <- unique( x )
            
            y <- sort( unlist( y ) )
            
            y <-
              paste(
                y,
                collapse = ", "
              )
              
            return( y )
            
          }
      )
    }
  )

table1 <- t( simplify2array( x = data.table1, higher = FALSE ) )

```

```{r }

names.included <-
  c(
    "Author.Year.Journal",
    "Population",
    "Age",
    "Time from diagnosis",
    "Contrast",
    "N.Limiting.Median",
    "Source",
    "N.Lipids.Reported"
  )

table1.printed <-
  table1[ , names.included ]

table1.printed <-
  as.data.frame(
    table1.printed,
    check.names = FALSE,
    stringsAsFactors = FALSE
  )

table1.printed$"Independence" <-
  round( 
    x = independence.authors[ table1.printed$"Author.Year.Journal" ] * 100
  )

colnames( table1.printed ) <-
  c(
    "Study",
    "Country",
    "Age",
    "Time from diagnosis",
    "Contrast",
    "Limiting sample size (median)",
    "Source of results",
    "Number of associated lipids reported",
    "Independence from other studies (%)"
  )

table1.printed[ table1.printed == "NA" ] <- "-"

# View( table1.printed )
# knitr::kable( table1.printed )
# knitr::kable( tibble::as_tibble( table1.printed ) )

# print(
#   x = tibble::as_tibble( table1.printed ),
#   n = Inf
#   # ,
#   # width = Inf
# )

write.table(
  x = table1.printed,
  file = "results-updated/Summary-Publications.tsv",
  row.names = FALSE,
  sep = "\t"
)

```

\newpage

# Number of Reported Lipids by Sample Size

```{r Significant-by-Sample-Size, message = FALSE, warning = FALSE}

data.plot <- data[ !is.na( data$"N.Limiting" ), ]

data.plot[ is.na( data.plot$"Age" ), "Age" ] <- "NA"

data.plot$"Time from diagnosis.filled" <- data.plot$"Time from diagnosis"

data.plot[ 
  is.na( data.plot$"Time from diagnosis.filled" ),
  "Time from diagnosis.filled"
  ] <- "NA"

data.plot$"Paper.Contrast.Age.Time" <-
  paste(
    data.plot$"Author.Year.Journal",
    data.plot$"Contrast",
    data.plot$"Age",
    data.plot$"Time from diagnosis.filled"
  )

data.plot.summary <- 
  by(
    data = data.plot,
    INDICES = data.plot$"Paper.Contrast.Age.Time",
    FUN = function( x ) {
      c(
        x[ 1, "Author.Year.Journal" ],
        x[ which( !is.na( x$"Significance" ) )[ 1 ], "Significance" ],
        any( !is.na( x$"Time from diagnosis" ) ),
        length( which( !is.na( x$"Significance" ) ) ),
        median( x$"N.Limiting", na.rm = TRUE )
      )
    }
  )

data.plot.summary <- t( simplify2array( x = data.plot.summary ) )

data.plot.summary <- 
  data.frame( 
    data.plot.summary,
    stringsAsFactors = FALSE
    )

colnames( data.plot.summary ) <-
  c(
    "Author.Year.Journal",
    "Significance",
    "is_post_diagnosis",
    "Markers_Reported",
    "Limiting_N"
  )

data.plot.summary$"is_multiple_testing_corrected" <- 
  grepl(
    x = data.plot.summary$"Significance",
    pattern = "adj"
  )

data.plot.summary$"Markers_Reported" <- 
  as.numeric( data.plot.summary$"Markers_Reported" )

data.plot.summary$"Limiting_N" <- as.numeric( data.plot.summary$"Limiting_N" )

data.plot.summary$"is_post_diagnosis" <- 
  unlist( data.plot.summary$"is_post_diagnosis" )

data.plot.summary$"is_multiple_testing_corrected_and_is_post_diagnosis" <-
  paste(
    data.plot.summary$"is_multiple_testing_corrected",
    data.plot.summary$"is_post_diagnosis"
  )

ggplot2::ggplot(
  data = data.plot.summary,
  mapping =
    ggplot2::aes(
      x = Limiting_N,
      y = Markers_Reported,
      color = is_multiple_testing_corrected_and_is_post_diagnosis
    )
) +
  ggplot2::geom_point() +
  ggplot2::geom_smooth( method = "lm", se = FALSE )

data.plot.summary.pre <-
  data.plot.summary[
    data.plot.summary$"is_post_diagnosis" == FALSE &
      data.plot.summary$is_multiple_testing_corrected == FALSE,
  ]

ggplot2::ggplot(
  data = data.plot.summary.pre,
  mapping =
    ggplot2::aes(
      x = Limiting_N,
      y = Markers_Reported
    )
) +
  ggplot2::geom_point() +
  ggplot2::geom_smooth( method = "auto", se = TRUE )

```

# Appendix

```{r SessionInfo}

utils::sessionInfo()

```